---
title: "ISD exponent derivation"
author: "Jeff Wesner"
date: '2022-10-28'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
```

Reuman et al. (2008, Ecology Letters) derive the ISD exponent from metabolic theory. Based on their equation 7 on page 1220, ISD exponents are explained by:

$$
b + 1 = \frac{log_{10}\alpha}{log_{10}\beta} -\frac{3}{4}
$$

where $\alpha$ is the trophic transfer efficiency, ranging from 0 (no energy transfer) to 1 (all energy is transferred) and $\beta$ is the body mass ratio of consumers to their resources. $\beta$ is assumed to be \>\>1, since predators are assumed to be much larger than their prey. The 3/4 is derived from the quarter scaling laws.

Across a broad range of $\alpha$ and $\beta$, these are the prediction of the ISD exponent.

```{r}
alpha = seq(0.01,1, length.out = 10)
beta = seq(5, 500, length.out = 50)


simulations = tibble(alpha = alpha) %>% 
  expand_grid(beta = beta) %>% 
  mutate(b = (log(alpha)/log(beta)) - 3/4 - 1)


simulations %>% 
  ggplot(aes(x = beta, y = b, color = alpha)) + 
  geom_line(aes(group = alpha)) +
  ylim(-5, 0) +
  geom_hline(yintercept = -1.5) +
  labs(title = "ISD predictions assuming the 3/4 power scaling")
```

As shown in the figure above, predicted values of the ISD exponent converge to \~2 at moderate levels of trophic transfer efficiency ($\alpha$) and at large ratios of consumer to resource sizes ($\beta$). Our samples with NEON are typically closer to -1.5 or less. However, from this equation, the only way to generate samples close to that empirical value is to change the -3/4 scaling assumption, which implies changing the -3/4 value. For example, changing -3/4 to -1/3 places the predicted values closer to the observed levels.

```{r}
alpha = seq(0.01,1, length.out = 10)
beta = seq(5, 500, length.out = 50)


simulations = tibble(alpha = alpha) %>% 
  expand_grid(beta = beta) %>% 
  mutate(b = (log(alpha)/log(beta)) - 1/3 - 1)


simulations %>% 
  ggplot(aes(x = beta, y = b, color = alpha)) + 
  geom_line(aes(group = alpha)) +
  ylim(-5, 0) +
  geom_hline(yintercept = -1.5) +
  labs(title = "ISD predictions assuming the 3/4 power scaling")
```
